---
params:
  ghDate: "`r Sys.Date() - 2`" # default to ~curent versions
title: "Base TEmplate for\n `r params$UnitName`"
author:
  - name: "Tom Philippi" 
    affiliation: IMD Species Inventory Program
  - name:  "Lisa Nelson"
    affiliation: IMD Species Inventory Program
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
---

```{r setup, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#-- Global settings
options(timeout = 600)
gc() # Call a garbage collection to free up memory

 
#-- Packages
# R packages used in this script
pkgLst <- c("dplyr",       # data management
            "report",      # automated citations
            "glue",        # text string editing
            "ggplot2",     # plotting/graphing
            "here",        # navigating directories
            "kableExtra",  # table formatting
            "knitr",       # Markdown formatting
            "rmarkdown",   # for render()
            "distill",     # for extended YAML
            "lubridate",   # dating
            "readr",       # writing TSV files
            "tibble",      # data structures
            "tidyr")       # data management

#-- Install
if (!"groundhog" %in% installed.packages()) install.packages("groundhog")

# n.b.: groundhog.install() does not like the current date as the date, 
# because CRANS update with uneven latency.  Therefore, the default value
# of params$ghDate is 2 days before the current date (when the script is running).
groundhog::groundhog.library(pkgLst, date = params$ghDate)

# create citations for R packages
# this is here for visibility in the template.  It should be re-run at the bottom
# in case additional packages are added along the way.
citelist <- report::cite_packages(include_R = TRUE)  #|>
#  report::clean_citation()


#-- knitr
# Set global options for RMarkdown
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      results = 'hide', 
                      comment = "",
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "Figures/",
                      tidy = TRUE, 
                      tidy.opts = list(width.cutoff = 60), 
                      cache = FALSE)


#-- ggplot2 theme
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(plot.title = ggplot2::element_text(hjust = 0.5))
sppPage <- 40


```

# Citations

`r report::clean_citation(report::cite_packages(include_R = TRUE)`

Kinseth, M and L. Nelson (2022).
  Boundary-derived Areas of Analysis for National Park Service Units, Fall 2021.
  NPS/NRSS/DRR---2022/3.
  <https://irma.nps.gov/DataStore/Reference/Profile/2287628>.

National Park Service Inventory and Monitoring Division (NPS-IMD; 2021).
  National Park Service NRSS Inventory and Monitoring Division (IMD) Management 
  Areas Inventory Data Services.
  National Park Service Data Store
  <https://irma.nps.gov/DataStore/Reference/Profile/2286496>.

National Park Service Inventory and Monitoring Division (NPS-IMD; 2022).
  NPS Unit Boundary-derived Areas of Analysis, Fall 2021.
  National Park Service Data Store.
  <https://irma.nps.gov/DataStore/Reference/Profile/2287631>.

R Core Team (2022). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria. URL
  <https://www.R-project.org/>.

# Session

This report was generated on `r lubridate::now()`. 
R session information is printed below.

```{r session, results='hold'}
sessionInfo()
```
